Fast Bayesian Inference for Exoplanet Discovery in Radial Velocity Data

### 2014-10-26 Paper/ms.tex
Posterior for N (traditional way)

\section{Inference}
Bayesian inference is the use of probability theory to describe uncertainty
\citep{sivia, ohagan}. In this framework, we approach data analysis problems by first
constructing a {\it hypothesis space}, which is the set of possible answers
to the problem we are considering. Normally, this is the set of possible
values of a vector of parameters $\btheta$ whose values we want
to know. We then assign probability distributions
called the {\it prior} and the {\it sampling distribution}. The prior
distribution $p(\btheta)$ describes
our initial uncertainty about which values of the parameters
$\btheta$ are plausible, and the
sampling distribution $p(\bdata | \btheta)$ describes
our initial uncertainty about the data set we're going to observe, as a
function of the unknown parameters $\btheta$.
When the data is known, our state of knowledge about the parameter is
updated from the prior $p(\theta)$ to the posterior distribution given by
Bayes' rule:
\begin{eqnarray}
p(\btheta | \bdata) &=&
\frac{p(\btheta)p(\bdata | \btheta)}
{p(\bdata)}
\end{eqnarray}
where the denominator, often called the {\it evidence} or
{\it marginal likelihood}, is given by:
\begin{eqnarray}
\mathcal{Z} = p(\bdata) = \int p(\btheta) p(\bdata | \btheta) d^n \btheta
\end{eqnarray}
where the integral is over the entire $n$-dimensional parameter space.

The number of orbiting planets, $N$, is an important parameter.
To calculate the posterior distribution for $N$, most authors
consider various trial values of $N$, and calculate the marginal likelihood
\begin{eqnarray}
p(\bdata | N) &=& \int p(\btheta | N) p(\bdata | \btheta, N) \, d^n \theta
\end{eqnarray}
for each possible value of $N$
\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}. Popular
methods for calculating the marginal likelihood are Nested Sampling
\citep{skilling} and ideas related to thermodynamic integration
\citep[e.g.][]{neal}.

>>>>

\section{Inference}
Bayesian inference is the use of probability theory to describe uncertainty
\citep{sivia, ohagan}. In this framework, we approach data analysis problems by first
constructing a {\it hypothesis space}, which is the set of possible answers
to the problem we are considering. Normally, this is the set of possible
values of a vector of parameters $\btheta$ whose values we want
to know. We then assign probability distributions
called the {\it prior} and the {\it sampling distribution}. The prior
distribution $p(\btheta)$ describes
our initial uncertainty about which values of the parameters
$\btheta$ are plausible, and the
sampling distribution $p(\bdata | \btheta)$ describes
our initial uncertainty about the data set we're going to observe, as a
function of the unknown parameters $\btheta$.
When the data is known, our state of knowledge about the parameter is
updated from the prior $p(\theta)$ to the posterior distribution given by
Bayes' rule:
\begin{eqnarray}
p(\btheta | \bdata) &=&
\frac{p(\btheta)p(\bdata | \btheta)}
{p(\bdata)}
\end{eqnarray}
where $p(\bdata | \btheta)$ is called the likelihood, once the actual dataset
is substituted in.
The denominator, often called the {\it evidence} or
{\it marginal likelihood}, is given by the expected value of the likelihood
with respect to the prior:
\begin{eqnarray}
\mathcal{Z} = p(\bdata) = \int p(\btheta) p(\bdata | \btheta) d^n \btheta
\end{eqnarray}
where the integral is over the entire $n$-dimensional parameter space.

The number of orbiting planets, $N$, is an important parameter.
To calculate the posterior distribution for $N$, most authors
consider various trial values of $N$, and calculate the marginal likelihood
\begin{eqnarray}
p(\bdata | N) &=& \int p(\btheta | N) p(\bdata | \btheta, N) \, d^n \theta
\end{eqnarray}
for each possible value of $N$
\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}.
The posterior distribution for $N$ can then be found straightforwardly by
using Bayes' rule with $N$ as the only unknown parameter:
\begin{eqnarray}
p(N | \bdata) &=& \frac{p(N)p(\bdata | N)}{\sum_N p(N)p(\bdata | N)}.
\end{eqnarray}
Popular
methods for calculating the marginal likelihood are Nested Sampling
\citep{skilling} and ideas related to thermodynamic integration
\citep[e.g.][]{neal}.
